---
title: "Homework 2"
author: "Ryan Wang"
date: "12/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tibble)
library(tidyverse)
```
# Bayesian Statistical Modeling Winter 2023
# Homework, Week 2

*When is homework due?* Homework is due on Mondays (Tuesday if Monday is a holiday). This assignment is due on Monday, 1/23/2022. Submit homework to gradescope: 

*Before submitting your work* You should use this file as a template for completing the assignment. You should change the author to your name. You should use the features of Rmarkdown to write your text answers below and include R code and equations where necessary. You should turn in both the .Rmd file as well as the pdf file that you generate by using the `Knit` button.


Complete problems from the end of chapter 2.

## 2E1

The probability of rain on Monday = given it is Monday, what it is probability of rainning.
2) and 4).


## 2E2

3).

## 2M1

1) WWW
```{r}
posterior_table1 =tibble(p_w_proposed = seq(from=0, to = 1, by = 0.01) )
posterior_table1 <- mutate(posterior_table1 , prior=1/n())
posterior_table1 <- mutate(posterior_table1 , likelihood=dbinom(3,size=3,prob=p_w_proposed))
posterior_table1 <- mutate(posterior_table1 , raw_posterior = prior*likelihood)
posterior_table1 <- mutate(posterior_table1 , posterior = raw_posterior/sum(raw_posterior))

ggplot(data=posterior_table1, aes(x=p_w_proposed, y=posterior))+
  geom_line(color="red")+
  geom_line(aes(x=p_w_proposed, y=prior) , color="blue")+
  labs(x = "probability of water", y = "posterior probability")
```

2) WWWL
```{r}
posterior_table2 =tibble(p_w_proposed = seq(from=0, to = 1, by = 0.01) )
posterior_table2 <- mutate(posterior_table2 , prior=1/n())
posterior_table2 <- mutate(posterior_table2 , likelihood=dbinom(3,size=4,prob=p_w_proposed))
posterior_table2 <- mutate(posterior_table2 , raw_posterior = prior*likelihood)
posterior_table2 <- mutate(posterior_table2 , posterior = raw_posterior/sum(raw_posterior))

ggplot(data=posterior_table2, aes(x=p_w_proposed, y=posterior))+
  geom_line(color="red")+
  geom_line(aes(x=p_w_proposed, y=prior) , color="blue")+
  labs(x = "probability of water", y = "posterior probability")


# Or we can update what we have from posterior_table1
d_now = 0
posterior_table22 <- mutate(posterior_table1 , likelihood=(d_now==1)*p_w_proposed + (d_now==0)*(1-p_w_proposed))
posterior_table22 <- mutate(posterior_table22 , raw_posterior = posterior*likelihood)
posterior_table22 <- mutate(posterior_table22 , posterior = raw_posterior/sum(raw_posterior))

ggplot(data=posterior_table22, aes(x=p_w_proposed, y=posterior))+
  geom_line(color="red")+
  geom_line(aes(x=p_w_proposed, y=prior) , color="blue")+
  labs(x = "probability of water", y = "posterior probability")
```

3) LWW LWWW
```{r}
posterior_table3 =tibble(p_w_proposed = seq(from=0, to = 1, by = 0.01) )
posterior_table3 <- mutate(posterior_table3 , prior=1/n())
posterior_table3 <- mutate(posterior_table3 , likelihood=dbinom(5,size=7,prob=p_w_proposed))
posterior_table3 <- mutate(posterior_table3 , raw_posterior = prior*likelihood)
posterior_table3 <- mutate(posterior_table3 , posterior = raw_posterior/sum(raw_posterior))

ggplot(data=posterior_table3, aes(x=p_w_proposed, y=posterior))+
  geom_line(color="red")+
  geom_line(aes(x=p_w_proposed, y=prior) , color="blue")+
  labs(x = "probability of water", y = "posterior probability")
```



## 2M2
1) WWW
```{r}
posterior_table4 =tibble(p_w_proposed = seq(from=0, to = 1, by = 0.01) )
posterior_table4 <- mutate(posterior_table4 , prior=ifelse(p_w_proposed < 0.5, 0, 2/(n()+1)))
sum(posterior_table4$prior)

posterior_table4 <- mutate(posterior_table4 , likelihood=dbinom(3,size=3,prob=p_w_proposed))
posterior_table4 <- mutate(posterior_table4 , raw_posterior = prior*likelihood)
posterior_table4 <- mutate(posterior_table4 , posterior = raw_posterior/sum(raw_posterior))

ggplot(data=posterior_table4, aes(x=p_w_proposed, y=posterior))+
  geom_line(color="red")+
  geom_line(aes(x=p_w_proposed, y=prior) , color="blue")+
  labs(x = "probability of water", y = "posterior probability")
```
2) WWWL
```{r}
posterior_table4 =tibble(p_w_proposed = seq(from=0, to = 1, by = 0.01) )
posterior_table4 <- mutate(posterior_table4 , prior=ifelse(p_w_proposed < 0.5, 0, 2/(n()+1)))
sum(posterior_table4$prior)

posterior_table4 <- mutate(posterior_table4 , likelihood=dbinom(3,size=4,prob=p_w_proposed))
posterior_table4 <- mutate(posterior_table4 , raw_posterior = prior*likelihood)
posterior_table4 <- mutate(posterior_table4 , posterior = raw_posterior/sum(raw_posterior))

ggplot(data=posterior_table4, aes(x=p_w_proposed, y=posterior))+
  geom_line(color="red")+
  geom_line(aes(x=p_w_proposed, y=prior) , color="blue")+
  labs(x = "probability of water", y = "posterior probability")
```

3) LWW LWWW
```{r}
posterior_table5 =tibble(p_w_proposed = seq(from=0, to = 1, by = 0.01) )
posterior_table5 <- mutate(posterior_table5 , prior=ifelse(p_w_proposed < 0.5, 0, 2/(n()+1)))
sum(posterior_table5$prior)

posterior_table5 <- mutate(posterior_table5 , likelihood=dbinom(5,size=7,prob=p_w_proposed))
posterior_table5 <- mutate(posterior_table5 , raw_posterior = prior*likelihood)
posterior_table5 <- mutate(posterior_table5 , posterior = raw_posterior/sum(raw_posterior))

ggplot(data=posterior_table5, aes(x=p_w_proposed, y=posterior))+
  geom_line(color="red")+
  geom_line(aes(x=p_w_proposed, y=prior) , color="blue")+
  labs(x = "probability of water", y = "posterior probability")
```

## 2M3

Pr(water|Earth) = 70%
Pr(land|Earth) = 1-70% =30%

Pr(land|Mars) = 1
Pr(Earth) = Pr(Mars) = 50%

Pr(Earth|land) = Pr(Earth and land) / Pr(land) 
= [Pr(land|Earth) * Pr(Earth)]/[Pr(land|Mars) * Pr(Mars) + Pr(land|Earth) * Pr(Earth)]
= (0.3 * 0.5) / (0.3 * 0.5 + 1 * 0.5) = 3/13 = 0.23

## 2H1

Pr(twins|A) = 0.1 
Pr(twins|B) = 0.2

Pr(A)=0.5
Pr(B)=0.5

Pr(twins and A) = Pr(twins|A) * Pr(A) = 0.1 * 0.5 = 0.05
Pr(twins and B) = Pr(twins|B) * Pr(B) = 0.2 * 0.5 = 0.1
Pr(twins)  = Pr(twins and B) + Pr(twins and A) = 0.15

Pr(A|twins) = Pr(twins|A) * Pr(A)/Pr(twins) = 0.1 * 0.5/ 0.15 = 1/3
Pr(B|twins) = Pr(twins|B) * Pr(B)/Pr(twins) = 0.2 * 0.5/ 0.15 = 2/3

Pr(Next_twins) = Pr(A|twins) * Pr(twins|A) + Pr(B|twins) * Pr(twins|B)
= 1/3 * 0.1 /+ 2/3 * 0.2 = 1/6



